Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-07-07 eventPointTwoLoop.tex
little clarification

\begin{equation}\label{eq:NLSMostrich}
\begin{aligned}
\text{Cut}\left(\! \scaleIntBscalarsmall{1}{2}{3}{4} \! \right)&=\sum_{\text{states}}\mathcal{A}^{\text{NLSM}}_{(2|q_4\bar{q}_3|\bar{q}_1)}\mathcal{A}^{\text{NLSM}}_{(1|\bar{q}_4q_3|\bar{q}_2)}\mathcal{A}^{\text{NLSM}}_{(q_2|34|q_1)}
\\
&=C^{\text{Xbox}}_{([12]|34)}\left[\tau^{(3)}_1 \tau^{(2)}_3\tau^{(1)}_4+\tau^{(3)}_1 \tau^{(2)}_4\tau^{(1)}_3+\tau^{(3)}_2 \tau^{(2)}_3\tau^{(1)}_4+\tau^{(3)}_2 \tau^{(2)}_4\tau^{(1)}_3\right]
\\
&-\left(C^{\text{2box}}_{(12|34)}\tau^{(3)}_1+C^{\text{2box}}_{(12|43)}\tau^{(3)}_2\right) \left[ \tau^{(2)}_3\tau^{(1)}_3+ \tau^{(2)}_4\tau^{(1)}_4\right] \,.
\end{aligned}
\end{equation}
The kinematic variables $\tau^{(i)}_j $ and $\tau_{ij}$ are the same, except we have made the replacement $p_i \rightarrow q_i$. The full two-loop NLSM amplitude can thus be computed by summing over the distinct labels of the resulting integrals, each weighted by suitable internal symmetry factors:
\begin{equation}
\label{eq:NLSM2loop}
\mathcal{A}^{\text{NLSM}}_{\text{2-loop}}=\frac{1}{4}\left[ \scaleIntCsmall \right]+ \frac{1}{2}\left[\scaleIntBscalarsmall{1}{2}{3}{4}+\scaleIntBscalarsmall{4}{3}{2}{1}\right]+\text{cyc}(2,3,4)
\end{equation}
We stress that the integrals appearing above are complicated by tensor integrals with many powers of loop momenta. However, as noted in the introduction, since both integrals are recursively one-loop, we can again apply a two-loop generalization of Passarino-Veltman for both the bubble and triangle integrals, stated in \eqn{eq:bubRed} and \eqn{eq:triRed}, respectively. With this, we proceed to the next step in the calculation.

@@

\begin{equation}\label{eq:NLSMostrich}

@@

We have added the additional factor of $(l_2+k_1)^2$ in the definition of $[I_3\circ I_2]^D(k_{12})$ to simplify the final expression in \eqn{eq:NLSMostrich}. We will use this as our basis integral throughout the text. This two-loop integral can be evaluated analytically using the $D$-dimensional single scale bubble and triangle integral expressions in \eqn{eq:integrals}. Thus, we have completely reduced the full two-loop NLSM amplitude to a linear combination of scalar integrals, $I_2^D(k_{ij})$ and $[I_3\circ I_2]^D(k_{ij})$. Now we will proceed to the final step of evaluating the amplitude in particular dimensions.
\paragraph{\textbf{Integration}} While there is plenty of physics to be extracted from the combination of \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we will focus on the behavior of the leading divergence for $D=2-2\epsilon$ in the planar limit where $N_c \rightarrow \infty$. In this limit, the cross-box is subleading in the $N_c$ expansion, and can be neglected. In a trace basis, the double-box can be expanded as follows:

@@

The analogous divergences for $D=4-2\epsilon$ can be extracted directly from the integral expressions in \eqn{eq:integrals}. Plugging these values into \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we can extract the full leading logarithmic divergence from the full color dressed amplitude in \eqn{eq:NLSM2loop}. Since we are interested in the large $N_c$ limit, we will define a color-ordered quantity $A^{2-2\epsilon}_{(1234)}$, 

@@

As in the previous section, exposed legs implicitly sum over internal states. The grey blobs above represent $D$-dimensional $t_8F^4$ operator insertions from the four-point BI tree amplitudes, which we labelled as $\mathcal{M}^{\gamma\gamma\gamma\gamma}_{(1234)}$. We give the internal loop momenta the same internal labels as we did the pions in \eqn{eq:NLSM2Bub} and \eqn{eq:NLSMostrich}, giving us the following integrals to evaluate for two-loop pure BI, 

@@

We should point out an additional verification that can occur for any SUSY amount of DBIVA.  We can recycle multiloop SYM integrands to verify our cuts.  In general this is because double-copy works at tree-level, and we can always write the Jacobi-satisfying Yang-Mills cut numerators in terms of color-ordered sYM amplitudes, $A^{\text{sYM}}$, offering us a KLT between cut-integrands.  When we cut to four-point amplitudes, as we do for the cuts above, this is particularly simple as only one order needs contribute for each for-point tree, yielding:
\begin{align}
\text{Cut}({\text{DBIVA}})&= \sum_\text{states} \prod_i \mathcal{A}^{\text{DBIVA}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\,\\
&= \sum_\text{states} \prod_i  \left( s^{(i)}_{12} s^{(i)}_{23}  A^{\text{sYM}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\right)\,\\
&= \left( \prod_j  \left(s^{(j)}_{12} s^{(j)}_{23}\right)  \right) \sum_\text{states}   \prod_i  A^{\text{sYM}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\,\\
&= \left( \prod_j  \left(s^{(j)}_{12} s^{(j)}_{23}\right) \right) \times \text{Cut}({\text{sYM}})\,.
\end{align}
Here  the products run over the number of trees contributing to the cut,  $s^{(i)}_{mn} = ( p^{(i)}_m + p^{(i)}_n )^2$, and $p^{(i)}_m$ is the $m$-th label of the $i$-th tree contributing to the cut. The second line follows from the double-copy structure of four-point amplitudes in DBIVA, where:
\begin{align}
\mathcal{A}_4^{\text{DBIVA}} = s_{12} \, s_{23} \, A^{\text{sYM}}(1,2,3,4)\,.
\end{align}

@@

\subsubsection{Two-loop $\mathcal{N}=4$ DBIVA} \label{sec:DBIvDC2loop}
At two-loop it sufficient to show that replacing the color factors reproduces the integrals we found above with the unitarity construction. Starting with the NLSM integrands in \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich}, we make the same replacement of color factors post integration as was done at one-loop.  We carry this out first for the double-bubble, and find that the full $D$-dimensional NLSM integral simplifies dramatically:
\begin{equation}
\begin{aligned}
\scaleIntCsmall \bigg|^{C_g \rightarrow N^{\mathcal{N}=4}_g}  \!\!&= N^{(12|34)}_{\mathcal{N}=4}\left[\frac{(s_{12}I_2^D(k_{12}))^2}{2}\left(\frac{s_{14}-s_{13}}{(D-1)^2}+s_{12}\right)\right]
\\
&+N^{(12|43)}_{\mathcal{N}=4}\left[\frac{(s_{12}I_2^D(k_{12}))^2}{2}\left(\frac{s_{13}-s_{14}}{(D-1)^2}+s_{12}\right)\right]
\\
&=s_{12}^4 (I_2^{D}(k_{12}))^2(t_8 F^4)^{\text{(max)}}_{(1234)}   \equiv \scaleIntCtune{}{hgreen}{}{hgreen} 
\end{aligned}
\end{equation}
The cancellation between dimension dependent factors is even more startling for the ostrich-diagram integral:
\begin{equation}
\begin{aligned}
\scaleIntBscalarsmall{1}{2}{3}{4}\bigg|^{C_g \rightarrow N^{\mathcal{N}=4}_g} \!\!&= N^{(12|34)}_{\mathcal{N}=4}\frac{s_{12}}{6} \left[\frac{(D-1)(D-4)s_{14}+2(D-2)^2 s_{13}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})
\\
&+N^{(12|34)}_{\mathcal{N}=4}\frac{s_{12}}{6} \left[\frac{(D-1)(D-4)s_{13}+2(D-2)^2 s_{14}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})
\\
&+N^{([12]|34)}_{\mathcal{N}=4}\left[\frac{s_{12}^2}{6}\frac{D+1}{D-1}\right][I_3\circ I_2]^D(k_{12})
\\
&=\frac{s_{12}^3}{3}[I_3\circ I_2]^{D}(k_{12})(t_8 F^4)^{\text{(max)}}_{(1234)} \equiv \scaleIntBtune{}{}{}{}{hgreen}{hgreen}{hgreen}{hgreen} 
\end{aligned}
\end{equation}
This alone is sufficient to demonstrate the validity of the double-copy, as these integrated quantities are exactly the same as those produced via generalized unitarity of $\mathcal{N}=4$ DBIVA at two-loop in \eqn{eq:N42bub} and \eqn{eq:N4ostrich}. We emphasize that based on our analysis, this consistency of the double-copy at two loop appears to hold in {any} dimension, as all the $D$-dependent prefactors drop out upon replacing the gauge theory color factors with $\mathcal{N}=4$ numerators. Considering the consistency of these two construction, this also serves as strong evidence for the existence of color-dual representation for two-loop pion integrands. We leave identifying such valid representation as an enticing future direction worth investigation.

@@

\paragraph{\textbf{Exponentiation}} In \eqn{eq:NLSM2bub} and \eqn{eq:NLSMostrich} we computed in general dimension, $D$, the two contributions to NLSM two-loop amplitudes. Evaluating these diagrams in $D=2-2\epsilon$, and taking the planar limit, $N_c \rightarrow \infty$, we found in \eqn{eq:expNLSM} that the IR divergences exponentiate through two-loop. This non-trivial property is found in theories that are conjectured to be integrable \cite{Shankar:1977cm,Zamolodchikov:1977nu,Zamolodchikov:1978xm, Komatsu:2019hgc}, like $\mathcal{N}=4$ super-Yang-Mills \cite{Anastasiou:2003kj,Bern:2005iz}. In future work, we hope to study whether this iterative structure can be further applied to the scale dependent logarithms present at multi-loop order in the theory. 

>>>>

\begin{equation}\label{eq:NLSMDunce}
\begin{aligned}
\text{Cut}\left(\! \scaleIntBscalarsmall{1}{2}{3}{4} \! \right)&=\sum_{\text{states}}\mathcal{A}^{\text{NLSM}}_{(2|q_4\bar{q}_3|\bar{q}_1)}\mathcal{A}^{\text{NLSM}}_{(1|\bar{q}_4q_3|\bar{q}_2)}\mathcal{A}^{\text{NLSM}}_{(q_2|34|q_1)}
\\
&=C^{\text{Xbox}}_{([12]|34)}\left[\tau^{(3)}_1 \tau^{(2)}_3\tau^{(1)}_4+\tau^{(3)}_1 \tau^{(2)}_4\tau^{(1)}_3+\tau^{(3)}_2 \tau^{(2)}_3\tau^{(1)}_4+\tau^{(3)}_2 \tau^{(2)}_4\tau^{(1)}_3\right]
\\
&-\left(C^{\text{2box}}_{(12|34)}\tau^{(3)}_1+C^{\text{2box}}_{(12|43)}\tau^{(3)}_2\right) \left[ \tau^{(2)}_3\tau^{(1)}_3+ \tau^{(2)}_4\tau^{(1)}_4\right] \,.
\end{aligned}
\end{equation}
The kinematic variables $\tau^{(i)}_j $ and $\tau_{ij}$ are the same, except we have made the replacement $p_i \rightarrow q_i$. The full two-loop NLSM amplitude can thus be computed by summing over the distinct labels of the resulting integrals, each weighted by suitable internal symmetry factors:
\begin{equation}
\label{eq:NLSM2loop}
\mathcal{A}^{\text{NLSM}}_{\text{2-loop}}=\frac{1}{4}\left[ \scaleIntCsmall \right]+ \frac{1}{2}\left[\scaleIntBscalarsmall{1}{2}{3}{4}+\scaleIntBscalarsmall{4}{3}{2}{1}\right]+\text{cyc}(2,3,4)
\end{equation}
We stress that the integrals appearing above are complicated by tensor integrals with many powers of loop momenta. However, as noted in the introduction, since both integrals are recursively one-loop, we can again apply a two-loop generalization of Passarino-Veltman for both the bubble and triangle integrals, stated in \eqn{eq:bubRed} and \eqn{eq:triRed}, respectively. With this, we proceed to the next step in the calculation.

@@

\begin{equation}\label{eq:NLSMdunce}

@@

We have added the additional factor of $(l_2+k_1)^2$ in the definition of $[I_3\circ I_2]^D(k_{12})$ to simplify the final expression in \eqn{eq:NLSMdunce}. We will use this as our basis integral throughout the text. This two-loop integral can be evaluated analytically using the $D$-dimensional single scale bubble and triangle integral expressions in \eqn{eq:integrals}. Thus, we have completely reduced the full two-loop NLSM amplitude to a linear combination of scalar integrals, $I_2^D(k_{ij})$ and $[I_3\circ I_2]^D(k_{ij})$. Now we will proceed to the final step of evaluating the amplitude in particular dimensions.
\paragraph{\textbf{Integration}} While there is plenty of physics to be extracted from the combination of \eqn{eq:NLSM2bub} and \eqn{eq:NLSMdunce}, we will focus on the behavior of the leading divergence for $D=2-2\epsilon$ in the planar limit where $N_c \rightarrow \infty$. In this limit, the cross-box is subleading in the $N_c$ expansion, and can be neglected. In a trace basis, the double-box can be expanded as follows:

@@

The analogous divergences for $D=4-2\epsilon$ can be extracted directly from the integral expressions in \eqn{eq:integrals}. Plugging these values into \eqn{eq:NLSM2bub} and \eqn{eq:NLSMdunce}, we can extract the full leading logarithmic divergence from the full color dressed amplitude in \eqn{eq:NLSM2loop}. Since we are interested in the large $N_c$ limit, we will define a color-ordered quantity $A^{2-2\epsilon}_{(1234)}$, 

@@

As in the previous section, exposed legs implicitly sum over internal states. The grey blobs above represent $D$-dimensional $t_8F^4$ operator insertions from the four-point BI tree amplitudes, which we labelled as $\mathcal{M}^{\gamma\gamma\gamma\gamma}_{(1234)}$. We give the internal loop momenta the same internal labels as we did the pions in \eqn{eq:NLSM2Bub} and \eqn{eq:NLSMDunce}, giving us the following integrals to evaluate for two-loop pure BI, 

@@

We should point out an additional verification that can occur for any SUSY amount of DBIVA.  We can recycle multiloop SYM integrands to verify our cuts.  In general this is because double-copy works at tree-level, and we can always write the Jacobi-satisfying Yang-Mills cut numerators in terms of color-ordered sYM amplitudes, $A^{\text{sYM}}$, offering us a KLT between cut-integrands.  When we cut to four-point amplitudes, as we do for the cuts above, this is particularly simple as only one order needs contribute for each for-point tree, yielding:
\begin{align}
\text{Cut}({\text{DBIVA}})&= \sum_\text{states} \prod_i \mathcal{A}^{\text{DBIVA}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\,\\
&= \sum_\text{states} \prod_i  \left( s^{(i)}_{12} s^{(i)}_{23}  A^{\text{sYM}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\right)\,\\
&= \left( \prod_j  \left(s^{(j)}_{12} s^{(j)}_{23}\right)  \right) \sum_\text{states}   \prod_i  A^{\text{sYM}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\,\\
&= \left( \prod_j  \left(s^{(j)}_{12} s^{(j)}_{23}\right) \right) \times \text{Cut}({\text{sYM}})\,.
\end{align}
Here  the products run over the number of trees contributing to the cut,  $s^{(i)}_{mn} = ( p^{(i)}_m + p^{(i)}_n )^2$, and $p^{(i)}_m$ is the $m$-th label of the $i$-th tree contributing to the cut. The second line follows from the double-copy structure of four-point amplitudes in DBIVA, where:
\begin{align}
\mathcal{A}_4^{\text{DBIVA}} = \mathcal{M}_4^{\text{NLSM} \otimes \text{sYM}}= s_{12} \, s_{23} \, A^{\text{sYM}}(1,2,3,4)\,.
\end{align}

@@

\subsubsection{Two-loop $\mathcal{N}=4$ DBIVA} \label{sec:DBIvDC2loop}
At two-loop it sufficient to show that replacing the color factors reproduces the integrals we found above with the unitarity construction. Starting with the NLSM integrands in \eqn{eq:NLSM2bub} and \eqn{eq:NLSMdunce}, we make the same replacement of color factors post integration as was done at one-loop.  We carry this out first for the double-bubble, and find that the full $D$-dimensional NLSM integral simplifies dramatically:
\begin{equation}
\begin{aligned}
\scaleIntCsmall \bigg|^{C_g \rightarrow N^{\mathcal{N}=4}_g}  \!\!&= N^{(12|34)}_{\mathcal{N}=4}\left[\frac{(s_{12}I_2^D(k_{12}))^2}{2}\left(\frac{s_{14}-s_{13}}{(D-1)^2}+s_{12}\right)\right]
\\
&+N^{(12|43)}_{\mathcal{N}=4}\left[\frac{(s_{12}I_2^D(k_{12}))^2}{2}\left(\frac{s_{13}-s_{14}}{(D-1)^2}+s_{12}\right)\right]
\\
&=s_{12}^4 (I_2^{D}(k_{12}))^2(t_8 F^4)^{\text{(max)}}_{(1234)}   \equiv \scaleIntCtune{}{hgreen}{}{hgreen} 
\end{aligned}
\end{equation}
The cancellation between dimension dependent factors is even more startling for the ostrich-diagram integral:
\begin{equation}
\begin{aligned}
\scaleIntBscalarsmall{1}{2}{3}{4}\bigg|^{C_g \rightarrow N^{\mathcal{N}=4}_g} \!\!&= N^{(12|34)}_{\mathcal{N}=4}\frac{s_{12}}{6} \left[\frac{(D-1)(D-4)s_{14}+2(D-2)^2 s_{13}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})
\\
&+N^{(12|34)}_{\mathcal{N}=4}\frac{s_{12}}{6} \left[\frac{(D-1)(D-4)s_{13}+2(D-2)^2 s_{14}}{(D-1)(4-3D)}\right][I_3\circ I_2]^D(k_{12})
\\
&+N^{([12]|34)}_{\mathcal{N}=4}\left[\frac{s_{12}^2}{6}\frac{D+1}{D-1}\right][I_3\circ I_2]^D(k_{12})
\\
&=\frac{s_{12}^3}{3}[I_3\circ I_2]^{D}(k_{12})(t_8 F^4)^{\text{(max)}}_{(1234)} \equiv \scaleIntBtune{}{}{}{}{hgreen}{hgreen}{hgreen}{hgreen} 
\end{aligned}
\end{equation}
This alone is sufficient to demonstrate the validity of the double-copy, as these integrated quantities are exactly the same as those produced via generalized unitarity of $\mathcal{N}=4$ DBIVA at two-loop in \eqn{eq:N42bub} and \eqn{eq:N4ostrich}. We emphasize that based on our analysis, this consistency of the double-copy at two loop appears to hold in {any} dimension, as all the $D$-dependent prefactors drop out upon replacing the gauge theory color factors with $\mathcal{N}=4$ numerators. Considering the consistency of these two construction, this also serves as strong evidence for the existence of color-dual representation for two-loop pion integrands. We leave identifying such valid representation as an enticing future direction worth investigation.

@@

\paragraph{\textbf{Exponentiation}} In \eqn{eq:NLSM2bub} and \eqn{eq:NLSMdunce} we computed in general dimension, $D$, the two contributions to NLSM two-loop amplitudes. Evaluating these diagrams in $D=2-2\epsilon$, and taking the planar limit, $N_c \rightarrow \infty$, we found in \eqn{eq:expNLSM} that the IR divergences exponentiate through two-loop. This non-trivial property is found in theories that are conjectured to be integrable \cite{Shankar:1977cm,Zamolodchikov:1977nu,Zamolodchikov:1978xm, Komatsu:2019hgc}, like $\mathcal{N}=4$ super-Yang-Mills \cite{Anastasiou:2003kj,Bern:2005iz}. In future work, we hope to study whether this iterative structure can be further applied to the scale dependent logarithms present at multi-loop order in the theory. 
